---
title: "SDM Tutorial - Jade"
output: html_document
date: "2023-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dismo

```{r}
install.packages("dimso")
install.packages("sp")
install.packages("raster")
install.packages("terra")
library("dismo")
help("dismo")
vignette('sdm','dismo')
```



## Green turtles

```{r}
library(readr)
Occurrence <- read_csv("~/Downloads/e1dfe463-237b-41ce-8103-adbf286404b1/Occurrence.csv")

o_cols <- c('decimallongitude','decimallatitude','date_year','scientificname','coordinateuncertaintyinmeters','bathymetry','sst','sss')

o_data <- data.frame(Occurrence)[o_cols]
good <- which(o_data$coordinateuncertaintyinmeters < 200)
o_data <- o_data[good,]
summary(o_data)
write.csv(o_data, file = "~/ohw23_proj_marinesdms/data/raw-bio/o_data.csv")

```

```{r}
install.packages('maptools')
install.packages('sp')
library(maptools)
data("wrld_simpl")
plot(wrld_simpl, xlim=c(30,90), ylim=c(-10,30), axes=TRUE, col="light yellow") +
# restore the box around the map
  box() +
# add the points
  points(o_data$decimallongitude, o_data$decimallatitude, col='orange', pch=20, cex=0.75) +
# plot points again to add a border, for better visibility
  points(o_data$decimallongitude, o_data$decimallatitude, col='red', cex=0.75)
```


# Training and testing data

```{r}
# presence train/test
set.seed(10)
group <- dismo::kfold(o_data, k = 5)  # 5 groups = 20 test/80 train split
pres_train <- o_data[group != 1, ]
pres_test <- o_data[group == 1, ]
```


# Trying out SDMpredictors with this particular data

```{r}
library(maxnet)
library(dplyr)
library(sf)
library(stars)
library(geodata)
library(dismo)
library(sdmpredictors)
```
Marine environmental datasets
```{r}
datasets <- list_datasets(terrestrial = FALSE, marine = TRUE)
layers <- list_layers(datasets)
View(layers)
```

Environmental data layers to use + cropping to the area of interest
```{r}
layercodes <- c("BO_sstmean","BO_salinity")
env <- load_layers(layercodes)

lons <- c(30, 90) 
lats <- c(-10,30)

ext <- raster::extent(lons[1], lons[2], lats[1], lats[2])
extent_polygon <- as(ext, "SpatialPolygons") %>% st_as_sf()
sf::st_crs(extent_polygon) <- 4326 # Not sure what this value 4326 means
AreaOfInterest <- raster::crop(env, extent(extent_polygon))
plot(AreaOfInterest)
```

```{r}
library(sdmpredictors)
devtools::install_github("zoonproject/zoon", force = TRUE)
library('zoon')
devtools::install_github("lifewatch/marinespeed", force = TRUE)
library('marinespeed')
```

```{r}
#Getting Turtle Occurence Data File in Correct Format for Zoon SDM:
pointsTurtles <- SpatialPoints(o_data[,c("decimallongitude", "decimallatitude")])
#pointsTurtles <- spTransform(pointsTurtles, CRSobj="+proj=longlat +ellps=GRS80") -> What was this?
occfile <- tempfile(fileext = ".csv")
write.csv(cbind(coordinates(pointsTurtles), value=1), occfile)

workflow(
  occurrence = LocalOccurrenceData(
    occfile, occurrenceType="presence",
    columns = c("decimallongitude", "decimallatitude", "date_year")), 
  covariate = LocalRaster(stack(AreaOfInterest)),
  process = OneHundredBackground(seed = 42), # just the page number of random numbers
  model = LogisticRegression,
  output = PrintMap)
```

## Looking up statistics and correlations for marine annual layers (almost exact copy from http://lifewatch.github.io/sdmpredictors/)
```{r}

# filter out monthly layers
layers_annual <- layers[is.na(layers$month),]
    
layers_annual_stats <- layer_stats(layers_annual)
View(layers_annual_stats)
    
layers_annual_correlations <- layers_correlation(layers_annual)
View(layers_annual_correlations)
    
# create groups of layers where no layers in one group have a correlation > 0.7 with a layer from another group
layers_annual_groups <- correlation_groups(layers_annual_correlations, max_correlation=0.7)
    
# inspect groups ; heatmap plot for larger groups (if gplots library is installed)
install.packages("gplots")
library(gplots)    
for(group in layers_annual_groups) {
      group_correlation <- as.matrix(layers_annual_correlations[group, group, drop=FALSE])
      if(require(gplots) && length(group) > 4){
        heatmap.2(abs(group_correlation)
                 ,main = "Correlation"
                 ,col = "rainbow"      
                 ,notecol="black"      # change font color of cell labels to black
                 ,density.info="none"  # turns off density plot inside color legend
                 ,trace="none"         # turns off trace lines inside the heat map
                 ,margins = c(12,9)    # widens margins around plot
                 )
      } else {
        print(group_correlation)
      }
    }
```


##  Create SDM: Note that this requires the ZOON, ggplot2, cowplot and marinespeed packages to be installed.(NOT DONE)

```{r}
# Load equal area rasters and crop with the extent of the Baltic Sea
library(sdmpredictors)
library(zoon)
library(raster)
library(ggplot2)
library(cowplot)
library(marinespeed)

layercodes_2 <- c("MS_biogeo05_dist_shore_5m", "MS_bathy_5m", "BO_sstrange", "BO_sstmean", "BO_salinity")
env_2 <- load_layers(layercodes_2, rasterstack = FALSE)
env_3 <- stack(env_2)
ext <- raster::extent(lons[1], lons[2], lats[1], lats[2])
extent_polygon <- as(ext, "SpatialPolygons") %>% st_as_sf()
sf::st_crs(extent_polygon) <- 4326 # crs -> reference system number WGS84 
studyarea <- raster::crop(env_3, extent(extent_polygon))
plot(studyarea)




    # Compare statistics between the original and the Australian bathymetry
   # View(rbind(layer_stats("MS_bathy_5m"),calculate_statistics("Bathymetry Australia", raster(australia, layer = 2))))

    # Compare correlations between predictors, globally and for Australia
    prettynames <- list(BO_salinity="Salinity", BO_sstmean="SST (mean)", 
                        BO_sstrange="SST (range)", MS_bathy_5m="Bathymetry",
                        MS_biogeo05_dist_shore_5m = "Shore distance")
    
   # p1 <- plot_corr(layers_correlation(layercodes), prettynames)
   # australian_correlations <- pearson_correlation_matrix(australia)
   # p2 <- plot_correlation(australian_correlations, prettynames)
   # cowplot::plot_grid(p1, p2, labels=c("A", "B"), ncol = 2, nrow = 1)
   # print(correlation_groups(australian_correlations))
    
# Fetch occurrences and prepare for ZOON
   occ <- marinespeed::get_occurrences(o_data)
    points <- SpatialPoints(occ[,c("longitude", "latitude")],
                            lonlatproj)
    points <- spTransform(points, equalareaproj)
    occfile <- tempfile(fileext = ".csv")
    write.csv(cbind(coordinates(points), value=1), occfile)
    # Create SDM with ZOON
    workflow(
      occurrence = LocalOccurrenceData(
        o_data, occurrenceType="presence",
        columns = c("decimallongitude", "decimallatitude")), 
      covariate = LocalRaster(studyarea),
      process = OneHundredBackground(seed = 42),
      model = LogisticRegression,
      output = PrintMap)
    # Layer citations
    print(layer_citations(layercodes_2))
```



## Maxnet with stars
```{r}
suppressPackageStartupMessages({
  library(maxnet)
  library(dplyr)
  library(maxnet)
  library(sf)
  library(stars)
  library(geodata)
  library(dismo)
  library(lubridate)
  library(sdmpredictors)
})
```

```{r}
library(sf)
lats <- c(-0.125, 32.125); lons <- c(41.875, 70.125)
# raster extent is defined by west lon, east lon, south lat, north lat
ext <- raster::extent(lons[1], lons[2], lats[1], lats[2])
extent_polygon <- as(ext, "SpatialPolygons") %>% st_as_sf()
# we need to assign a coordinate system; 4326 is the default for maps in sf
sf::st_crs(extent_polygon)<-4326
```

```{r}
library("rnaturalearth")
library("rnaturalearthdata")
world <- ne_countries(scale = "medium", returnclass = "sf")
```

```{r}
library(ggplot2)
library(sf)
ggplot(data = world) +
    geom_sf() + 
  geom_sf(data = extent_polygon, color = "red", fill=NA)
```

```{r}
spp <- "Chelonia mydas"
fil <- file.path(here::here(), "data", "raw-bio", "io-sea-turtles.csv")
occ <- read.csv(fil)
occ <- occ %>% subset(scientificName == spp)
```

```{r}
library(tidyverse)
colnames(occ)

occ <- occ %>% subset(bathymetry > 0 & 
                        shoredistance > 0 & 
                        coordinateUncertaintyInMeters < 200)
dim(occ)

occ$date <- as.Date(occ$eventDate)
occ.sf <- sf::st_as_sf(occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

```{r}
library(ggplot2)
library("ggspatial")
library("sf")
theme_set(theme_bw())
world <- st_make_valid(world)
world_points <- st_centroid(world)
world_points <- cbind(world, st_coordinates(st_centroid(world$geometry)))


plt <- ggplot(data = world) +
    geom_sf(fill= "antiquewhite") +
    geom_point(data = occ, aes(x=decimalLongitude, y=decimalLatitude), color = "red", size=0.1) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.15, "in"), pad_y = unit(0.25, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = lons, ylim = lats) +
    theme(panel.grid.major = element_line(color = gray(.5), linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))

sf_use_s2(FALSE)
plt + geom_text(data = world_points, aes(x=X, y=Y, label=name),
          color = "darkblue", size=2, check_overlap = TRUE) +
  xlab("longitude") + ylab("latitude") + 
  ggtitle(spp, subtitle = "occurences since 2000")

```

```{r}
path <- tempdir()
recent <- geodata::worldclim_global(var="bio", res=10, path = path) |>
  stars::st_as_stars() |>
  split()
names(recent) <- sprintf("bio%0.2i", seq_len(length(names(recent))))
```


dataframe for the occurrence+env
```{r}
env_occ <- stars::st_extract(recent, sf::st_coordinates(occ.sf)) |>
  dplyr::as_tibble() |>
  na.omit()
env_occ
```

dataframe for the absence(background)+env
```{r}
back <- read_csv("~/ohw23_proj_marinesdms/data/raw-bio/pts_absence.csv")
colnames(back) <- c("decimallongitude","decimallatitude")
back.sf <- sf::st_as_sf(back,coords= c("decimallongitude","decimallatitude"), crs = 4326)

env_back <- stars::st_extract(recent, sf::st_coordinates(back.sf)) |>
  dplyr::as_tibble() |>
  na.omit()
env_back

```

Build the model. Create a flag vector that distinguishes between occurrence (1) and background (0).
```{r}
pres <- c(rep(1, nrow(env_occ)), rep(0, nrow(env_back)))
model <- maxnet::maxnet(pres,
                    dplyr::bind_rows(env_occ, env_back))
```

