---
title: "Get background data"
---

We used the random samples within our region of interest to generate a file with the locations for our absences.

## Set up

```{r}
library(ggplot2)
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")
library(raster)
library(tidyverse)
```

```{r}
dir_data <- file.path(here::here(), "data", "raw-bio")
dir_env <- file.path(here::here(), "data", "env")
redo <- TRUE
```

## Define the region of interest

```{r}
# lon, lon, lat, lat
library(sf)
lats <- c(-0.125, 32.125)
lons <- c(41.875, 65.125)
# raster extent is defined by west lon, east lon, south lat, north lat
ext <- raster::extent(lons[1], lons[2], lats[1], lats[2])
extent_polygon <- as(ext, "SpatialPolygons") %>% st_as_sf()
# we need to assign a coordinate system; 4326 is the default for maps in sf
sf::st_crs(extent_polygon)<-4326
# convert to a WKT format
wkt_geometry <- extent_polygon$geometry %>% st_as_text()
wkt_geometry
```

Make a map of our region so we know we have the right area.
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) + geom_sf() +
  geom_sf(data = extent_polygon, color = "red", fill=NA)
```

## Random samples

This is adapted from [here](https://bbest.github.io/eds232-ml/lab1a_sdm-explore.html).

### Get a marine raster layer

We just need one because we use this to sample lat/lons from the marine environment.

```{r}
# set a default data directory
options(sdmpredictors_datadir = dir_env)

# choosing marine
env_datasets <- sdmpredictors::list_datasets(terrestrial = FALSE, marine = TRUE)
env_layers <- sdmpredictors::list_layers("MARSPEC")
env_stack <- sdmpredictors::load_layers("MS_bathy_5m")
env_stack <- env_stack %>% raster::crop(ext)
```

Plot to check that the layer looks ok. This is bathymetry.

```{r}
plot(env_stack)
```
  
### Next we sample points from this


```{r}
nsamp <- 1000
absence <- dismo::randomPoints(env_stack[[1]], nsamp) %>% 
    as_tibble() %>% 
    st_as_sf(coords = c("x", "y"), crs = 4326)
```

```{r}
mapview::mapview(absence, col.regions = "gray")
```

Save the absence locations to a file.

```{r save-absence}
absence_geo <- file.path(dir_data, "absence.geojson")
pts_absence_csv <- file.path(dir_data, "pts_absence.csv")
st_write(absence, pts_absence_csv, layer_options = "GEOMETRY=AS_XY", append=FALSE)
```

