---
title: "Maxent SDM attempt"
output: html_document
date: "2023-08-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pre-SDM
### Set-up

You will need to install a new version of maxnet for this tutorial.
```{r, eval=FALSE}
devtools::install_github("BigelowLab/maxnet")
```

<<<<<<< HEAD
```{r, warning=F}
=======
```{r, warning=FALSE}
suppressPackageStartupMessages({
>>>>>>> 24641da453d589319595340e48eca999159b9d22
library(maxnet)
library(dplyr)
library(maxnet)
library(sf)
library(stars)
library(geodata)
library(dismo)
library(lubridate)
library(sdmpredictors)
library(zoon)
library(ggplot2)
library(cmocean)
library(janitor)
library(DT)
})
```

### Load data in

```{r, warning=F}
# presence data
io.turtles <- read.csv("/home/jovyan/R/ohw23_proj_marinesdms/data/raw-bio/io-sea-turtles.csv") 

# absence data
pts.abs <- read.csv("/home/jovyan/R/ohw23_proj_marinesdms/data/raw-bio/pts_absence.csv") # X is lon and Y is lat
```

### Format occurrence data

```{r, warning=F}
spp <- c("Chelonia mydas", "Caretta caretta", "Eretmochelys imbricata", "Lepidochelys olivacea", "Natator depressus", "Dermochelys coriacea") # turtle species we're interested in

occ <- io.turtles %>% 
  subset(scientificName == spp) # subsetting all the occurence data to just those turtles 

table(occ$scientificName) # seeing how often each species occurs

occ <- occ %>% # but we also need to subset the occurences to include just those in the water (not the ones on land)
  subset(bathymetry > 0 & 
                        shoredistance > 0 & 
                        coordinateUncertaintyInMeters < 200)

table(occ$scientificName) # seeing how often each species occurs now
# Caretta caretta is Loggerhead and Chelonia mydas is Green sea turtles 

occ.sub <- occ[,c(2,4,5,6,9,11,13,14)] # choosing the cols I want

occ.sub$eventDate <- lubridate::ymd_hms(occ.sub$eventDate) # changing to datetime format

colnames(occ.sub) <- c("sci.name", "obsv.datetime", "lat", "lon", "life.stage", "bathy", "SST", "SSS") # renaming
```

```{r, warning=F}
# format background/absence points

colnames(pts.abs) <- c("lon","lat")
pts.abs <- na.omit(pts.abs)

abs.points <- sf::st_as_sf(pts.abs, coords = c("lon", "lat"), crs = 4326)
```

```{r, warning=FALSE}
# format prescence/occurrence points

head(occ.sub)

occ.points <- sf::st_as_sf(occ.sub, coords = c("lon", "lat"), crs = 4326)
head(occ.points)

occ.points <- occ.points %>% 
  mutate(common.name = case_when(sci.name == "Caretta caretta" ~ "Loggerhead",
                                 sci.name == "Chelonia mydas" ~ "Green"))
```

### SDM predictors

Loading in

```{r, warning=F}
datasets <- sdmpredictors::list_datasets(terrestrial = FALSE, marine = TRUE)

layers <- list_layers(datasets)
#View(layers)
```

Choosing and formatting

```{r, warning=FALSE}
layercodes = c("BO_sstmean", "BO_bathymean", "BO22_ph", "BO2_dissoxmean_bdmean", "BO2_salinitymean_ss", "BO2_chlomean_ss", "BO21_nitratemean_ss") # the env variables I chose from SDMpredictors

env <- load_layers(layercodes, rasterstack = T) # take out the equalarea arg or you will be sad AND add rasterstack = T
env.stars <- stars::st_as_stars(env) # convert to stars object
env.stars <- split(env.stars)

lats <- c(-0.125, 32.125); lons <- c(41.875, 70.125) # IO lat/lon range
# raster extent is defined by west lon, east lon, south lat, north lat

ext <- raster::extent(lons[1], lons[2], lats[1], lats[2])
extent_polygon <- as(ext, "SpatialPolygons") %>% 
  st_as_sf()
# we need to assign a coordinate system; 4326 is the default for maps in sf
sf::st_crs(extent_polygon) <- 4326 # applying a coordinate system

plot(extent_polygon) # look a special rectangle

bb <- sf::st_bbox(extent_polygon) # make a bounding box 

# plot(env.stars["BO_bathymean"] %>% sf::st_crop(bb)) # looking at bathymetry

env.obs <- stars::st_extract(env.stars,
                             sf::st_coordinates(occ.points)) %>%
  dplyr::as_tibble()

io.rast <- raster::crop(env, extent(extent_polygon))
plot(io.rast) # look NOT a rectangle- your env variables cropped to the Indian Ocean!
```

### Pull in background data

```{r, warning=F}
env.back <- stars::st_extract(env.stars, sf::st_coordinates(abs.points)) %>% 
  dplyr::as_tibble() %>% 
  na.omit()

head(env.back)
```

## SDM Model

### Running model

```{r, warning=F}
env.obs <- na.omit(env.obs); env.back <- na.omit(env.back) # remove NA values

pres <- c(rep(1, nrow(env.obs)), rep(0, nrow(env.back))) # create values of 1 for presence data and 0 for absence data

sdm.model <- maxnet::maxnet(pres, rbind(env.obs, env.back))
```

### Model metrics

```{r, warning=F}
responses <- plot(sdm.model, type = "cloglog")
```

## Predicting

```{r, warning=F}
clamp <- TRUE       # see ?predict.maxnet for details
type <- "cloglog"
predicted <- predict(sdm.model, env.stars %>% sf::st_crop(bb), clamp = clamp, type = type)
predicted
```

## Visualization

```{r, warning=F}
# ggplot - with occurrence data points

ggplot() +
  geom_stars(data = predicted) +
  scale_fill_cmocean(name = "ice", direction = -1, guide = guide_colorbar(barwidth = 1, barheight = 10, ticks = FALSE, nbin = 1000, frame.colour = "black"), limits = c(0, 1)) +
  theme_linedraw() +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "Loggerhead and green sea turtle SDM in the Arabian Sea",
       x = "Longitude",
       y = "Latitude",
       fill = "Probability",
       shape = "Species (presence)",
       subtitle = "Environmental predictors: mean SS temp, mean SS salinity, mean bathymetry, \nmean pH, mean DO, mean SS chlorophyll-a, mean SS nitrate") +
  geom_point(occ.points, mapping = aes(shape = common.name, geometry = geometry), stat = "sf_coordinates", alpha = 0.3, color = "purple") +
  #scale_x_continuous(breaks = seq(40, 70, 10), limits = c(42, 70))+
  scale_y_continuous(breaks = seq(0, 30, 10))

# ggsave("SDM_loggerhead_green_w points.pdf", height = 6, width = 8.5)
```

```{r, warning=FALSE}
# ggplot - without occurrence data points

ggplot() +
  geom_stars(data = predicted) +
  scale_fill_cmocean(name = "ice", direction = -1, guide = guide_colorbar(barwidth = 1, barheight = 10, ticks = FALSE, nbin = 1000, frame.colour = "black"), limits = c(0, 1)) +
  theme_linedraw() +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "Loggerhead and green sea turtle SDM in the Arabian Sea",
       x = "Longitude",
       y = "Latitude",
       fill = "Probability",
       shape = "Species (presence)",
       subtitle = "Environmental predictors: mean SS temp, mean SS salinity, mean bathymetry,\nmean pH, mean DO, mean SS chlorophyll-a, mean SS nitrate") +
  #geom_point(occ.points, mapping = aes(shape = common.name, geometry = geometry), stat = "sf_coordinates", alpha = 0.3, color = "purple") +
  #scale_x_continuous(breaks = seq(40, 70, 10), limits = c(42, 70))+
  scale_y_continuous(breaks = seq(0, 30, 10))

# ggsave("SDM_loggerhead_green.pdf", height = 6, width = 8.5)
```


```{r, warning=F}
<<<<<<< HEAD
# ggplot - with occurrence (purple) and absence (green) data points
=======
# ggplot - with occurrence and absence data points
>>>>>>> 24641da453d589319595340e48eca999159b9d22

ggplot() +
  geom_stars(data = predicted) +
  scale_fill_cmocean(name = "ice", direction = -1, guide = guide_colorbar(barwidth = 1, barheight = 10, ticks = FALSE, nbin = 1000, frame.colour = "black"), limits = c(0, 1)) +
  theme_linedraw() +
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "Loggerhead and green sea turtle SDM in the Arabian Sea",
       x = "Longitude",
       y = "Latitude",
       fill = "Probability",
       shape = "Species (presence)",
       subtitle = "Environmental predictors: mean SS temp, mean SS salinity, mean bathymetry, \nmean pH, mean DO, mean SS chlorophyll-a, mean SS nitrate") +
  geom_point(occ.points, mapping = aes(shape = common.name, geometry = geometry), stat = "sf_coordinates", alpha = 0.3, color = "purple") +
  #scale_x_continuous(breaks = seq(40, 70, 10), limits = c(42, 70))+
  scale_y_continuous(breaks = seq(0, 30, 10)) +
  geom_point(abs.points, mapping = aes(geometry = geometry), stat = "sf_coordinates", alpha = 0.3, color = "green") # adding in absence data
```

