---
title: "Marine Species Distribution Models"
subtitle: "Maxent example"
date: 2023-08-07
output: html_document
---

This example steps through the process of presence-only modeling using the maxnet package. The example is based on the notes by Ben Tupper (Biglow Lab, Maine) https://github.com/BigelowLab/maxnet/wiki/stars

### Workflow

1. Define the area of interest
2. Obtain sea turtle occurrences for the NE coast of Africa
3. Use the sdmpredictors package to assemble predictor variable data
4.  Collect background points within the region occupied by the presence points using sf package
5. Model and predict using the maxnet package

### Load necessary packages

```{r}
suppressPackageStartupMessages({
  library(maxnet)
  library(dplyr)
  library(maxnet)
  library(sf)
  library(stars)
  library(geodata)
  library(dismo)
  library(lubridate)
  library(sdmpredictors)
})
```

## Step 1: Define the region of interest

We are looking at the western Arabian Sea, Persian Gulf, Gulf of Oman, Gulf of Aden and Red Sea.

```{r}
library(sf)
lats <- c(-0.125, 32.125); lons <- c(41.875, 70.125)
# raster extent is defined by west lon, east lon, south lat, north lat
ext <- raster::extent(lons[1], lons[2], lats[1], lats[2])
extent_polygon <- as(ext, "SpatialPolygons") %>% st_as_sf()
# we need to assign a coordinate system; 4326 is the default for maps in sf
sf::st_crs(extent_polygon)<-4326
```

Make a map of our region with the polygon of interest.
```{r}
library(ggplot2)
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")
ggplot(data = world) + geom_sf() +
  geom_sf(data = extent_polygon, color = "red", fill=NA)
```


## Step 2: Load the occurrence data

This was queried using the robis package.

```
spp <- c("Chelonia mydas", "Caretta caretta", "Eretmochelys imbricata", "Lepidochelys olivacea", "Natator depressus", "Dermochelys coriacea")
wkt_geometry <- extent_polygon$geometry %>% st_as_text()
df <- occurrence(spp, startdate = as.Date("2000-01-01"), geometry = wkt_geometry)
```

```{r}
spp <- "Chelonia mydas"
occ <- read.csv("/home/jovyan/R/ohw23_proj_marinesdms/data/raw-bio/io-sea-turtles.csv")
occ <- occ %>% subset(scientificName == spp)
```

Look at the column names
```{r}
library(tidyverse)
colnames(occ)
```

Make sure the turtle was at sea.
```{r}
occ <- occ %>% subset(bathymetry > 0 & 
                        shoredistance > 0 & 
                        coordinateUncertaintyInMeters < 200)
dim(occ)
```

Change the occurrence data into an sf object with a coordinate system. Add a date column with YYYY-MM-DD format.

```{r}
occ$date <- as.Date(occ$eventDate)
occ.sf <- sf::st_as_sf(occ, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

Plot the occurrence data

```{r}
library(ggplot2)
library("ggspatial")
library("sf")
theme_set(theme_bw())
world <- st_make_valid(world)
world_points <- st_centroid(world)
world_points <- cbind(world, st_coordinates(st_centroid(world$geometry)))


plt <- ggplot(data = world) +
    geom_sf(fill= "antiquewhite") +
    geom_point(data = occ, aes(x=decimalLongitude, y=decimalLatitude), color = "red", size=0.1) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.15, "in"), pad_y = unit(0.25, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = lons, ylim = lats) +
    theme(panel.grid.major = element_line(color = gray(.5), linetype = "dashed", size = 0.5), panel.background = element_rect(fill = "aliceblue"))

sf_use_s2(FALSE)
plt + geom_text(data = world_points, aes(x=X, y=Y, label=name),
          color = "darkblue", size=2, check_overlap = TRUE) +
  xlab("longitude") + ylab("latitude") + 
  ggtitle(spp, subtitle = "occurences since 2000")

```

## Use the sdmpredictors package to assemble predictor variable data

## Collect background points within the region occupied by the presence points using sf package

## Model and predict using the maxnet package

