---
title: "Get environmental data for the occurrence locationa"
---

We need to get a data frame with the environmental data for the occurrence locations.

## Set up

```{r}
library(ggplot2)
library(sdmpredictors)
```

Set the directory where we will save environmental data layers.

```{r}
dir_env <- here::here("data", "env")
options(sdmpredictors_datadir = dir_env)
```

## Add datasets

We will use the sdmpredictors R package which has marine data layers.

```{r}
# choose marine
env_datasets <- sdmpredictors::list_datasets(terrestrial = FALSE, marine = TRUE)
```

## Show the available variables

The dataframe is large. We will use the DT package to make the table pretty in html.
```{r}
env_layers <- sdmpredictors::list_layers(env_datasets$dataset_code)
DT::datatable(env_layers)
```

## Variables


#Loading in environmental marine data
datasets <- list_datasets(terrestrial = FALSE, marine = TRUE)
layers <- list_layers(datasets)
#View(layers)
layercodes <- c("BO_sstmean","BO_salinity")
env <- load_layers(layercodes)
