---
title: "get turtle data"
output: html_document
date: "2023-08-08"
---

Here we download from OBIS using the robis package.

## Step 1 define the region of interest

```{r}
# lon, lon, lat, lat
library(sf)
lats <- c(-0.125, 32.125)
lons <- c(41.875, 65.125)
# raster extent is defined by west lon, east lon, south lat, north lat
ext <- raster::extent(lons[1], lons[2], lats[1], lats[2])
extent_polygon <- as(ext, "SpatialPolygons") %>% st_as_sf()
# we need to assign a coordinate system; 4326 is the default for maps in sf
sf::st_crs(extent_polygon)<-4326
# convert to a WKT format
wkt_geometry <- extent_polygon$geometry %>% st_as_text()
wkt_geometry
```

Make a map of our region so we know we have the right area.
```{r}
library(ggplot2)
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) + geom_sf() +
  geom_sf(data = extent_polygon, color = "red", fill=NA)
```

## Get occurrence data from robis

```{r}
library(robis)
spp <- "Caretta caretta" # loggerhead
df <- occurrence(spp, startdate = as.Date("2000-01-01"), geometry = wkt_geometry)
```

Subset columns and save.
```{r}
cols.to.use <- c("occurrenceID", "scientificName", "dateIdentified", "eventDate", "decimalLatitude", "decimalLongitude", "coordinateUncertaintyInMeters",
                 "individualCount","lifeStage", "sex",
                 "bathymetry",  "shoredistance", "sst", "sss")
df <- df[,cols.to.use]
```

Save the data to a file for later use. The `here()` function is used so that we don't have to worry about where you have this Rmarkdown file loaded.
```{r eval = FALSE}
fil <- here::here("data", "raw-bio", "loggerhead-robis.csv")
write.csv(df, file=fil, quote=FALSE)
```


## Download data for four species

We will download data for four sea turtles found in the Arabian sea and save to one file.

```{r eval = FALSE}
spp <- c("Chelonia mydas", "Caretta caretta", "Eretmochelys imbricata", "Lepidochelys olivacea", "Natator depressus", "Dermochelys coriacea")
obs <- occurrence(spp, startdate = as.Date("2000-01-01"), geometry = wkt_geometry)
cols.to.use <- c("occurrenceID", "scientificName", "dateIdentified", "eventDate", "decimalLatitude", "decimalLongitude", "coordinateUncertaintyInMeters",
                 "individualCount","lifeStage", "sex",
                 "bathymetry",  "shoredistance", "sst", "sss")
obs <- obs[,cols.to.use]
```

## Save our data

Set up the file names
```{r}
dir_data <- here::here("data", "raw-bio")
filname <- "io-sea-turtles"
obs_csv <- file.path(dir_data, paste0(filname, ".csv"))
obs_geo <- file.path(dir_data, paste0(filname, ".geojson"))
obs_gpkg <- file.path(dir_data, paste0(filname, ".gpkg"))
```

Change the data frame to a sf dataframe.
```{r}
obs_sf <- obs %>% 
    sf::st_as_sf(
      coords = c("decimalLongitude", "decimalLatitude"),
      crs = st_crs(4326))
```

Save files.
```{r}
redo   <- TRUE

if (!file.exists(obs_csv) | redo)  readr::write_csv(obs, obs_csv)
if (!file.exists(obs_geo) | redo)  sf::write_sf(obs_sf, obs_geo, delete_dsn=TRUE)
if (!file.exists(obs_gpkg) | redo)  sf::write_sf(obs_sf, obs_gpkg, delete_dsn=TRUE)
```

We can reload our data as
```{r}
obs_sf2 <- sf::read_sf(obs_gpkg)
class(obs_sf2)
```

