---
title: "Fit SDM"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will use the presence and absence data and the environmental layers that we created in the previous steps to assemble the data frames needed to fit our SDM model.

## Pre-SDM Set-up

You will need to install a new version of maxnet for this tutorial. No need to do this on the JupyterHub. It is already installed.
```{r, eval=FALSE}
devtools::install_github("BigelowLab/maxnet")
install.packages(c("dplyr", "sf", "stars", "geodata",
                   "dismo", "lubridate", "sdmpredictors", 
                   "ggplot2", "cmocean", "janitor", "DT",
                   "here"))
```

Load the needed packages.
```{r, warning=FALSE}
suppressPackageStartupMessages({
library(maxnet)
library(dplyr)
library(sf)
library(stars)
library(geodata)
library(dismo)
library(lubridate)
library(sdmpredictors)
#library(zoon)
library(ggplot2)
library(cmocean)
library(janitor)
library(DT)
})
```

Set the file location.
```{r}
here::i_am("tutorial/Steps_sdm_maxnet.Rmd")
```

Load region files. We created our region objects in a separate file [Region data](./tutorial/00_Region.html) and saved these in `data/region`. We will load these now.

Load the bounding box polygon and create a bounding box.
```{r}
#Loading bounding box for the area of interest
fil <- here::here("data", "region", "BoundingBox.shp")
extent_polygon <- sf::read_sf(fil)
bbox <- sf::st_bbox(extent_polygon)
```

## Load occurence data

Here we load the data prepared in the previous step.

```{r, warning=F}
# presence data
fil <- here::here("data", "raw-bio", "io-sea-turtles-clean.csv")
occ.sub <- read.csv(fil)
```

Create sf_points data frame.

```{r, warning=FALSE}
occ.points <- sf::st_as_sf(occ.sub, coords = c("lon", "lat"), crs = 4326)
head(occ.points)
```

## Load background data

```{r, warning=F}
# absence data
fil <- here::here("data", "raw-bio", "pts_absence.csv")
pts.abs <- read.csv(fil) # X is lon and Y is lat
```

Check column names.

```{r}
colnames(pts.abs)
```

Change columns names to lon and lat and remove any NAs in the data.
```{r, warning=F}
colnames(pts.abs) <- c("lon","lat")
pts.abs <- na.omit(pts.abs)
```

Convert to sf_points object. Set the crs to 4326.
```{r}
abs.points <- sf::st_as_sf(pts.abs, coords = c("lon", "lat"), crs = 4326)
```

## Get the environment for the lat/lon locations

Here we create the data frame with the environmental variables for our presence and absence locations.

### Load environmental layers

Set the location of the data directory.
```{r}
dir_env <- here::here("data", "env")
options(sdmpredictors_datadir = dir_env)
```

Specify the layers that we want. The layers were saved to the `data/env` directory.
```{r, warning=FALSE}
layercodes <- c("BO_sstmean", "BO_bathymean", "BO22_ph", "BO2_dissoxmean_bdmean", "BO2_salinitymean_ss", "BO2_chlomean_ss", "BO21_nitratemean_ss")
```

Load the layers into the `env` object. We want to set rasterstack equal true to get one file for our variables.
```{r}
env <- sdmpredictors::load_layers(layercodes, rasterstack = TRUE)
```

### Create a stars object

There are a few ways that we can get the values of the environmental variables in our raster stack for the lat/lon locations. We will use the stars and terra functions as these are the new (2023) packages in R for this purpose. You will find older approaches if you search and possibly AI will suggest older approaches.

Step one is to convert our raster stack to a stars object and split the stars object into layers.

```{r}
env.stars <- stars::st_as_stars(env) # convert to stars object
env.stars <- terra::split(env.stars)
```

### Extract the variables for our points

Now we can extract the variables for our presence and absence points.
```{r}
occ.env <- stars::st_extract(env.stars, sf::st_coordinates(occ.points)) %>%
  dplyr::as_tibble()

abs.env <- stars::st_extract(env.stars, sf::st_coordinates(abs.points)) %>% 
  dplyr::as_tibble()
```

Now we have a data frame with the variables for presence and absence.
```{r}
head(abs.env)
```

Let's check the sizes. They should have the same number of rows. 
```{r}
dim(occ.points)
dim(occ.env)
```

Let's check the sizes. They should have the same number of rows. 
```{r}
dim(abs.points)
dim(abs.env)
```

Our environmental data have some NAs. We need to remove these.
```{r}
any(is.na(occ.env))
any(is.na(abs.env))
```

Remove the NAs.
```{r}
occ.env <- na.omit(occ.env)
abs.env <- na.omit(abs.env)
```

## SDM Model

We will fit with MaxEnt using the maxnet package. The function is
```
mod <- maxnet::maxnet(pres, env_df)
```
`env_df` is a data frame of environmental data (in columns) for each location (in rows) of the presences and absences. pres` is a string of 0s and 1s specifying if the row in `environ` is a presence (1) or absence (0).

### Set up the `env_df`

We combining the two data frames with `rbind()`.
```{r}
env_df <- rbind(occ.env, abs.env)
```

### Set up the `pres` string

The rows of `occ.env` are all 1 (presence) and the rows of `abs.env` are all 0 (absent).

```{r}
pres <- c(rep(1, nrow(occ.env)), rep(0, nrow(abs.env)))
```

## Run the model

```{r, warning=F}
sdm.model <- maxnet::maxnet(pres, env_df)
```

### Model metrics

```{r, warning=F}
responses <- plot(sdm.model, type = "cloglog")
```

## Save model

```{r, warning=F}
env.stars.crop <- env.stars %>% sf::st_crop(bbox)
fil <- here::here("data", "models", "io-turtle.RData")
save(sdm.model, env.stars.crop, occ.points, abs.points, file=fil)
```