---
title: "Get turtle data"
output: html_document
---

Here we download from OBIS using the robis package.

## Set up

```{r}
library(ggplot2)
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")
library(raster)
library(tidyverse)
library(robis)
```

## Load the region info

Load the bounding box polygon and create a bounding box.
```{r}
#Loading bounding box for the area of interest
fil <- here::here("data", "region", "BoundingBox.shp")
extent_polygon <- sf::read_sf(fil)
bbox <- sf::st_bbox(extent_polygon)
wkt_geometry <- extent_polygon$geometry %>% st_as_text()
```

Make a map of our region so we know we have the right area.
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) + geom_sf() +
  geom_sf(data = extent_polygon, color = "red", fill=NA)
```

## Get occurrence data from robis

We will download data for four sea turtles found in the Arabian sea and save to one file. We will use the `occurrence()` function in the robis package.

```{r eval = FALSE}
spp <- c("Chelonia mydas", "Caretta caretta", "Eretmochelys imbricata", "Lepidochelys olivacea", "Natator depressus", "Dermochelys coriacea")

obs <- robis::occurrence(spp, startdate = as.Date("2000-01-01"), geometry = wkt_geometry)
```

This has many columns that we don't need. We reduced to fewer columns.
```{r eval = FALSE}
cols.to.use <- c("occurrenceID", "scientificName", 
                 "dateIdentified", "eventDate", 
                 "decimalLatitude", "decimalLongitude", "coordinateUncertaintyInMeters",
                 "individualCount","lifeStage", "sex",
                 "bathymetry",  "shoredistance", "sst", "sss")
obs <- obs[,cols.to.use]
```

We also added a cleaner date with YYYY-MM-DD format.
```{r eval = FALSE}
obs$date <- as.Date(obs$eventDate)
```

## Save our data

Set up the file names
```{r eval = FALSE}
dir_data <- here::here("data", "raw-bio")
filname <- "io-sea-turtles"
obs_csv <- file.path(dir_data, paste0(filname, ".csv"))
obs_geo <- file.path(dir_data, paste0(filname, ".geojson"))
obs_gpkg <- file.path(dir_data, paste0(filname, ".gpkg"))
```

Change the data frame to a sf dataframe.
```{r eval = FALSE}
obs_sf <- obs %>% 
    sf::st_as_sf(
      coords = c("decimalLongitude", "decimalLatitude"),
      crs = st_crs(4326))
```

Save files in different formats to facilitate loading into geospatial packages.
```{r eval = FALSE}
redo   <- TRUE

if (!file.exists(obs_csv) | redo)  readr::write_csv(obs, obs_csv)
if (!file.exists(obs_geo) | redo)  sf::write_sf(obs_sf, obs_geo, delete_dsn=TRUE)
if (!file.exists(obs_gpkg) | redo)  sf::write_sf(obs_sf, obs_gpkg, delete_dsn=TRUE)
```

Later we can reload our data as
```{r eval = FALSE}
tmp <- sf::read_sf(obs_gpkg)
class(tmp)
```

## Clean and prep data

Here we clean and prepare the data for our model and save to a new file name.

### Load data in

```{r, warning=F}
# presence data
fil <- here::here("data", "raw-bio", "io-sea-turtles.csv")
io.turtles <- read.csv(fil)
```

### Clean the data

Select species.
```{r, warning=F}
# turtle species we're interested in
spp <- c("Chelonia mydas", "Caretta caretta", "Eretmochelys imbricata", "Lepidochelys olivacea", "Natator depressus", "Dermochelys coriacea") 

# subsetting all the occurence data to just those turtles 
occ <- io.turtles %>% 
  subset(scientificName == spp) 

# subset the occurences to include just those in the water
occ <- occ %>% 
  subset(bathymetry > 0 & shoredistance > 0 & coordinateUncertaintyInMeters < 200)

# seeing how often each species occurs
table(occ$scientificName) 
```

After cleaning we discover that we only have loggerhead and green sea turtles.

### Select the needed columns

Select columns and add a common name column.

```{r}
colnames(occ)
```

We want these. The last two are sea surface temperature and salinity.
```{r}
cols <- c("scientificName", "eventDate", "decimalLatitude", "decimalLongitude", "lifeStage", "bathymetry", "sst", "sss")
```

Subset the columns.
```{r}
occ.sub <- occ %>% dplyr::select(all_of(cols))
```

Fix the event date to a date format.
```{r}
occ.sub$eventDate <- lubridate::ymd_hms(occ.sub$eventDate) 
```

Change the column names.
```{r}
colnames(occ.sub) <- c("sci.name", "obsv.datetime", "lat", "lon", "life.stage", "bathy", "SST", "SSS")
```

Add common.name column.
```{r}
occ.sub <- occ.sub %>% 
  mutate(common.name = case_when(sci.name == "Caretta caretta" ~ "Loggerhead",
                                 sci.name == "Chelonia mydas" ~ "Green"))
```

### Save the cleaned file

## Save our data

Set up the file names
```{r}
dir_data <- here::here("data", "raw-bio")
fil <- "io-sea-turtles-clean.csv"
obs_csv <- here::here(dir_data, fil)
if (!file.exists(obs_csv))  readr::write_csv(occ.sub, obs_csv)
```