---
title: "Saving Environment Data"
---

We are going to download environment data layers using sdmpredictors R package. We need to specify the directory where we will do this. We created the env directory in the data directory for this purpose.

Then we need to convert this to a format that will allow us to get the environmental variables at the lat/lon values of our observations and background points.

```{r set-layers-dir}
#Location of hte data directory
dir_env <- here::here("data", "env")
#Set default directory for environmental data
options(sdmpredictors_datadir = dir_env)
```

### Load layer info

Loading in the available data layers. We specify `marine = TRUE`.

```{r, warning=FALSE}
datasets <- sdmpredictors::list_datasets(terrestrial = FALSE, marine = TRUE)
layers <- list_layers(datasets)
#View(layers)
```

### Choose layers

See the [Background discussion](./tuturials/03_sdmpredictors-variables.Rmd) on how we decided on the environmental variables that we would use from the sdmpredictors R package.

```{r, warning=FALSE}
layercodes <- c("BO_sstmean", "BO_bathymean", "BO22_ph", "BO2_dissoxmean_bdmean", "BO2_salinitymean_ss", "BO2_chlomean_ss", "BO21_nitratemean_ss")
```

### Load layers

We want to set rasterstack equal true to get one file for our variables. This will save the files to `data/env` and we can load the files in later steps.
```{r}
env <- sdmpredictors::load_layers(layercodes, rasterstack = TRUE)
```

## Visualize the environment data

Load the bounding box.

```{r}
#Loading bounding box for the area of interest
fil <- here::here("data", "region", "BoundingBox.shp")
extent_polygon <- read_sf(fil)
```

Make a plot of all the layers.
```{r}
io.rast <- raster::crop(env, extent(extent_polygon))
plot(io.rast) 
```
