---
title: "Get environmental data for the occurrence locations"
---

We are going to download environment data layers using sdmpredictors R package. We need to specify the directory where we will do this. We created the env directory in the data directory for this purpose.

Ultimately we need to get a data frame with the environmental data for the occurrence and absence lat/lon locations.

## Set up

```{r}
library(ggplot2)
library(sdmpredictors)
library(DT)
library(sf)
library(raster)
```

Set the directory where we will save environmental data layers.

```{r}
dir_env <- here::here("data", "env")
options(sdmpredictors_datadir = dir_env)
```

## Add datasets

We will use the sdmpredictors R package which has marine data layers.

```{r}
# choose marine
env_datasets <- sdmpredictors::list_datasets(terrestrial = FALSE, marine = TRUE)
```

## Show the available variables

The dataframe is large. We will use the DT package to make the table pretty in html.
```{r}
env_layers <- sdmpredictors::list_layers(env_datasets$dataset_code)
DT::datatable(env_layers)
```

## Variables

See the [Background discussion](./tuturials/03_sdmpredictors-variables.Rmd) on how we decided on the environmental variables that we would use from the sdmpredictors R package.

```{r, warning=FALSE}
layercodes <- c("BO_sstmean", "BO_bathymean", "BO22_ph", "BO2_dissoxmean_bdmean", "BO2_salinitymean_ss", "BO2_chlomean_ss", "BO21_nitratemean_ss")
```

### Load layers

We want to set rasterstack equal true to get one file for our variables. This will save the files to `data/env` and we can load the files in later steps.
```{r}
env <- sdmpredictors::load_layers(layercodes, rasterstack = TRUE)
```

## Visualize the environment data

Load the bounding box.

```{r}
#Loading bounding box for the area of interest
fil <- here::here("data", "region", "BoundingBox.shp")
extent_polygon <- read_sf(fil)
```

Make a plot of all the layers cropped to our bounding box.
```{r}
io.rast <- raster::crop(env, raster::extent(extent_polygon))
plot(io.rast) 
```

## Environmental predictors for points

We will use the `stars` package to sample from our raster layers.

Load in our point data as data frames.
```{r, warning=FALSE}
# presence data
fil <- here::here("data", "raw-bio", "io-sea-turtles-clean.csv")
df.occ <- read.csv(fil) 

# absence data
fil <- here::here("data", "raw-bio", "pts_absence.csv")
df.abs <- read.csv(fil)
colnames(df.abs) <- c("lon", "lat")
```

Convert data frames to sf points objects. This is what `stars` needs.
```{r, warning=F}
df.abs <- na.omit(df.abs) # just in case
sf.abs <- sf::st_as_sf(df.abs, coords = c("lon", "lat"), crs = 4326)
sf.occ <- sf::st_as_sf(df.occ, coords = c("lon", "lat"), crs = 4326)
```

Convert the raster stack to a stars object.
```{r}
env.stars <- stars::st_as_stars(env) # convert to stars object
env.stars <- terra::split(env.stars)
```

Get environment values for the absence points. Each row in `sf.abs` is a row in `env.abs`.
```{r, warning=FALSE}
env.abs <- stars::st_extract(env.stars, sf::st_coordinates(sf.abs)) %>% 
  dplyr::as_tibble() %>% 
  na.omit()

head(env.abs)
```

Get environment values for the occurence points. Each row in `sf.occ` is a row in `env.occ`.
```{r, warning=FALSE}
env.occ <- stars::st_extract(env.stars, sf::st_coordinates(sf.occ)) %>% 
  dplyr::as_tibble() %>% 
  na.omit()

head(env.occ)
```

Now make this into one data frame with a `pa` column for 1 is a occurrence row and 0 if an absence row.

```{r}
pres <- c(rep(1, nrow(env.occ)), rep(0, nrow(env.abs)))
sdm_data <- data.frame(pa = pres, rbind(env.occ, env.abs))
head(sdm_data)
```

Save to a file. We will use for other models.
```{r}
fil <- here::here("data", "raw-bio", "sdm_data.csv")
write.csv(sdm_data, row.names = FALSE, file=fil)
```