---
title: "env_vars"
output: html_document
date: "2023-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Second approach. Use Python from R

From the terminal window, do
```
conda create -n xarray xarray netcdf4 zarr
```

Now we can use Python and xarray in R. We can use 
```{r}
library(reticulate)
use_condaenv("xarray")
xr = import("xarray")
xr$open_dataset("/home/jovyan/shared/marine-SDMs/INDIAN_OCEAN_025GRID_DAILY.zarr")
```

But I find it hard to remember reticulate syntax, so it is easier to just make a py script or some python text and run that:

```{r}
python_code <- "
import pandas as pd

data = {'Name': ['Alice', 'Bob', 'Charlie'],
        'Age': [25, 30, 22]}
df = pd.DataFrame(data)
"
py_run_string(python_code)
```

Now we can convert the pandas dataframe to an R dataframe:
```{r}
df <- py$df %>% as.data.frame()
df
```

The py$ operator (provided by reticulate) gives you access to the objects in the Python environment. 

## Extracting data from the zarr file

So now we can use Python to extract from the zarr file the env vars at the occurance date, lat, lon values.


```{r}
py_code <- "
import xarray as xr
ds = xr.open_zarr('/home/jovyan/shared/marine-SDMs/INDIAN_OCEAN_025GRID_DAILY.zarr/')
ds.sst.sel(time="2022-01-01", lat=[2,4], lon=[65, 66], method="nearest").to_pandas()
"
