---
title: "env_vars"
output: html_document
date: "2023-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# First approach

Using the zarr file, xarray and python and saved that to netcdf. Red the netcdf file into R. Only worked because we set a small bounding box so the netcdf file was only 5G. This approach is not scalable and loads a lot of data into R that we don't need. We only need the environment variables where the species occurances are.

## With raster package

```{r}
# load packages  
library(sf) 
library(ncdf4)
library(raster)
library(rasterVis)
library(RColorBrewer)
```

## Read netCDF

```{r}
# set path and filename
ncpath <- "/home/jovyan/shared/marine-SDMs/IO_subset.nc"
tmp_raster <- raster::brick(ncpath, varname="CHL")
tmp_raster
```

```{r}
# rasterVis plot
mapTheme <- rasterTheme(region = rev(brewer.pal(10, "RdBu")))
cutpts <- c(-50,-40,-30,-20,-10,0,10,20,30,40,50)
plt <- levelplot(subset(tmp_raster, 1), margin = F, at=cutpts, cuts=11, pretty=TRUE, par.settings = mapTheme)
plt
```



```{r}
# rasterVis plot
plot(subset(tmp_raster, 1))
```

## With the ncdf4 package

```{r}
ncdf.sub <- nc_open("/home/jovyan/shared/marine-SDMs/IO_subset.nc")
lat <- ncvar_get(ncdf.sub, "lat", verbose = F)
lon <- ncvar_get(ncdf.sub, "lon")
t <- ncvar_get(ncdf.sub, "time")
chl.array <- ncvar_get(ncdf.sub, "CHL")
fillvalue <- ncatt_get(ncdf.sub, "CHL", "_FillValue")
fillvalue
chl.array[chl.array == fillvalue$value] <- NA
chl.slice <- chl.array[, , 1]
r <- raster(t(chl.slice), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
plot(r)
```

```{r}
names(ncdf.sub$var)
```

