---
title: "Working with georeferenced observations"
output:
  md_document:
    variant: markdown_github
date: "2023-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal of the notebook
In this notebook, we replicate the beginning of the [tutorial](../tutorial/Get_seaturtle_data.Rmd) on accessing data from OBIS.
  
What is changed is how we store the data both in memory and in file storage. When using spatial data (like point data for turtle observations), it good coding practice to leverage all of the spatial-data handling that R, Python and most modern languages provide.  The idea is that the coder can focus on the attributes (aka "variables!" "columns!") associated with each observation (aka "features!" "records!" or "rows!).  The spatial data (aka "geometry!") is handled silently for you - woot! woot!


# Establish the directory heirarchy

# Loading relevant libraries
```{r libraries}
suppressPackageStartupMessages({
  #Dealing with spatial data
  library(sf)
  #Getting base maps
  library(rnaturalearth)
  #Access to OBIS
  library(robis)
  #Data manipulation and visualisation
  library(tidyverse)
  library(janitor)
})
```

# Creating a bounding box
We will use a bounding box for the region of our interest (Arabian Sea and the Bay of Bengal) to extract *C. caretta* data relevant to our study area.  
  
```{r bounding_box}
#We create a bounding box using minimum and maximum coordinate pairs
extent_polygon <- sf::st_bbox(c(xmin = 41.875, xmax = 65.125, 
                            ymax = -0.125, ymin = 32.125), 
                          #Assign reference system
                          crs = sf::st_crs(4326)) |>    # the new pipe only needs 2 keystrokes! 
  #Turn into sf object
  sf::st_as_sfc()

#Extract polygon geometry as text - needed for the OBIS query
pol_geometry <- sf::st_as_text(extent_polygon[[1]])
```
  
# Searching for sea turtle occurrence from OBIS
We will use the `robis` package to find observations of Loggerhead sea turtles (*C. caretta*) published in OBIS.  

```{r obis_search}
#Search OBIS for loggerhead observations from 2000
caretta_obs <- robis::occurrence("Caretta caretta", 
                          startdate = as.Date("2000-01-01"),
                          #Apply spatial constraint
                          geometry = pol_geometry,
                          #Include absence records if available
                          absence = "include")
dplyr::glimpse(caretta_obs)
```

```{r go_spatial}
caretta_obs = caretta_obs |>
  sf::st_as_sf(coords = c("decimalLongitude", "decimalLatitude"),
               crs = 4326)
caretta_obs
```


# Save the data in [Geopackage format](https://www.geopackage.org/)
```{r save_spatial}
caretta_obs = sf::write_sf(caretta_obs, here::here("data", "raw-bio", "io-sea-turtles.gpkg"))
```


# Read it back
```{r read_spatial}
x = sf::read_sf(here::here("data", "raw-bio", "io-sea-turtles.gpkg"))
x
```
