---
title: "Getting sea turtle observations"
output:
  md_document:
    variant: markdown_github
date: "2023-08-08"
---

# Goal of the notebook
In this notebook, we will show how you can get some [Loggerhead sea turtle (*Caretta caretta*)](https://www.iucnredlist.org/species/3897/119333622) data from 2000 until present from the [Ocean Biodiversity Information System (OBIS)](https://obis.org/). We will use the `robis` package to search the OBIS library and download relevant data.  
  
# Loading relevant libraries
```{r libraries, warning = F, message = F}
#Dealing with spatial data
library(sf)
#Getting base maps
library(rnaturalearth)
#Access to OBIS
library(robis)
#Data manipulation and visualisation
library(tidyverse)
library(janitor)
```

# Creating a bounding box
We will use a bounding box for the region of our interest (Arabian Sea and the Bay of Bengal) to extract *C. caretta* data relevant to our study area.  
  
```{r bounding_box}
#We create a bounding box using minimum and maximum coordinate pairs
extent_polygon <- st_bbox(c(xmin = 41.875, xmax = 65.125, 
                            ymax = -0.125, ymin = 32.125), 
                          #Assign reference system
                          crs = st_crs(4326)) %>% 
  #Turn into sf object
  st_as_sfc()

#Extract polygon geometry 
pol_geometry <- st_as_text(extent_polygon[[1]])
```
  
# Searching for sea turtle occurrence from OBIS
We will use the `robis` package to find observations of Loggerhead sea turtles (*C. caretta*) published in OBIS.  

```{r obis_search}
#Search OBIS for loggerhead observations from 2000
caretta_obs <- occurrence("Caretta caretta", 
                          startdate = as.Date("2000-01-01"),
                          #Apply spatial constraint
                          geometry = pol_geometry,
                          #Include absence records if available
                          absence = "include")

#Check structure of results
glimpse(caretta_obs)
```